Parsing Postfix log files

### 2007-12-14 doc/logparser.tex
Using pdflatex to generate a PDF directly works now. It needs some fine tuning: the graphs look a bit crap, and pdflatex complains     about the flow charts, but it's working. Also removed the range from the tables, because I now understand standard     deviation and I was using it incorrectly.
Improved the logs2regexs algorithm explanation.

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs; this doesn't work in the bibliography, but the breakurl
% package does.
%\usepackage{url}
% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\PLP{}}
\newcommand{\parsernames}{\PLP{}'s}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}

@@

% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author don't work - I think either gnuplot or graphviz clobbers it.
% With later TeX versions the dvips argument is necessary; everywhere else
% suggests using pdftex as the argument, but pdftex doesn't know how to
% include postscript images, so I need to create a dvi, convert it to
% postscript, and convert that to PDF.  I might lookup how to include
% postscript with pdftex.
\usepackage[dvips]{hyperref}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
%    breaklinks  = true,
}
% This is necessary for URLs in the bibliography; I dunno why the url
% package doesn't work.
\usepackage{breakurl}
\usepackage{lastpage}

@@

    \item [Calculate word frequencies]  The position of words within a line
        is important: a common word does not indicate similarity between
        lines unless it occupies the same position within both
        lines.\footnote{If a variable term within a line contains spaces,
        it will appear to the algorithm as two words rather than one.  This
        will alter the position of subsequent words, so a word occurring in
        different positions in two log lines may indicate similarity, but
        the algorithm does not attempt to deal with this possibility.}  The
        algorithm maintains a counter for each \textit{(word, word's
        position within the line)\/} tuple, incrementing it each time that
        word is occurs in that position.

@@

    \item [Classify words based on their frequency]  The frequency of each
        \textit{(word, word's position within the line)\/} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all lines is generally a good starting point) it
        is classified as a fixed word, otherwise it is classified as a
        variable term.  Variable terms are replaced by \texttt{.*}, which
        means to match zero or more of any character.  The words are
        reassembled to produce a \regex{} matching the line, and a counter
        is maintained for each \regex{}.

    \item [Build regexes]  Contiguous sequences of \texttt{.*} in the newly
        reassembled \regexes{} are collapsed to a single \texttt{.*}; any
        resulting duplicate \regexes{} are combined, and their counters
        summed.  If the frequency of a \regex{} is lower than the threshold
        supplied by the user the \regex{} is discarded.\footnote{This is a
        second, independent threshold, but once again 1\% of input lines is
        a good starting point; obviously the threshold depends on the
        number and type of input lines.}  The new \regexes{} are printed
        for the user to add to the database, either as new rules or merged
        into the \regexes{} of existing rules; the counter for each
        \regex{} is also printed, giving the user an indication of how many
        of the input lines that \regex{} should match.  Discarding
        \regexes{} will result in some of the input lines not being
        matched; this utility should be run again once the unmatched lines
        have been isolated by running the parser, including the new
        \regexes{}, with the same input which produced the original set of
        unparsed lines.

>>>>

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs; this doesn't work in the bibliography, but the breakurl
% package does.
%\usepackage{url}
% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\PLP{}}
\newcommand{\parsernames}{\PLP{}'s}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}
% Provides commands to distinguish between pdf and dvi output.
\usepackage{ifpdf}

@@

% When creating a \PDF{} make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author don't work - I think either gnuplot or graphviz clobbers it.
\ifpdf{}
    \usepackage[pdftex]{hyperref}
\else{}
    \usepackage[dvips]{hyperref}
\fi{}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
%    breaklinks  = true,
}
% This is necessary for URLs in the bibliography; I dunno why the url
% package doesn't work; don't include it when generating \PDF{} output.
\ifpdf{}
\else{}
    \usepackage{breakurl}
\fi{}
\usepackage{lastpage}

@@

    \item [Calculate word frequencies]  The position of words within a line
        is important: a common word does not indicate similarity between
        lines unless it occupies the same position within both
        lines.\footnote{If a variable term within a line contains spaces,
        it will appear to the algorithm as two words rather than one.  This
        will alter the position of subsequent words, so a word occurring in
        different positions in two log lines may indicate similarity, but
        the algorithm does not attempt to deal with this possibility.}  The
        algorithm maintains a counter for each \textit{(word, word's
        position within the line)\/} tuple, incrementing it each time that
        word occurs in that position.

@@

    \item [Classify words based on their frequency]  The frequency of each
        \textit{(word, word's position within the line)\/} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all lines is generally a good starting point) it
        is classified as a fixed word, otherwise it is classified as a
        variable term.  Variable terms are replaced by \texttt{.*}, which
        means to match zero or more of any character.

    \item [Build regexes]  The words are reassembled to produce a \regex{}
        matching the line, and a counter is maintained for each \regex{}.
        Contiguous sequences of \texttt{.*} in the newly reassembled
        \regexes{} are collapsed to a single \texttt{.*}; any resulting
        duplicate \regexes{} are combined, and their counters summed.  If
        the frequency of a \regex{} is lower than the threshold supplied by
        the user the \regex{} is discarded.\footnote{This is a second,
        independent threshold, but once again 1\% of input lines is a good
        starting point; obviously the threshold depends on the number and
        type of input lines.}  The new \regexes{} are printed for the user
        to add to the database, either as new rules or merged into the
        \regexes{} of existing rules; the counter for each \regex{} is also
        printed, giving the user an indication of how many of the input
        lines that \regex{} should match.  Discarding \regexes{} will
        result in some of the input lines not being matched; this utility
        should be run again once the unmatched lines have been isolated by
        running the parser, including the new \regexes{}, with the same
        input which produced the original set of unparsed lines.
